---
title: "Peer assesment 1"
output: html_document
---

I will first load my data 

```{r}
activity <- read.csv("~/Downloads/activity.csv")
```

I know have to make sure the data is in a proceed way so that my reults are adecuate.

```{r, echo=FALSE}
head(activity)
steps <- activity [,1]
date <- activity [,2]
interval <- activity [,3]
```

I first notice that there are a lot of Na missing values in the dataset.

I would like to make a lattice histogram to ilustrate the total number of steps taken each day.

```{r,}
hist(tapply(steps, date, sum), xlab = "Total daily steps", breaks = 20, 
    main = "Steps taken per day")
```

But I will like to calculate the mean and media of everyday to do so i will :

```{r,warning=FALSE}
library(stats)
sum <-  tapply(steps, date,sum)
mean(sum,na.rm = TRUE)
median(sum,na.rm = TRUE)
```
The question was:

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

-   here is a time serie

```{r,warning=FALSE}
ts <- ts(activity$interval,frequency=5)
library(lattice)
 xyplot(steps ~ ts, activity,
        type = "l",na.rm = TRUE)
```
 
 Noe lets find out in which interval there is the highest mean.
 
```{r}
mean_interval <- as.numeric(tapply(steps,activity$interval, mean, na.rm = TRUE))
interval_ordered <-  activity$interval[order(mean_interval, decreasing = TRUE)]
head(interval_ordered)
```
We can see that the maximun average steps is in the 835 interval
There are many missing values but how can i measure that?
```{r,warning=FALSE}
summary(activity)
## I can get there the missing vaulues but if i call the folowing function i will know the percentage.
 mean (is.na(activity))
```

Now lets fill those missing values with the 0 value

```{r,}
nactivity <- activity
head(activity)
nsteps <- nactivity [,1]
ndate <- nactivity [,2]
ninterval <-  nactivity [,3]
nactivity[is.na(nactivity)] <- 0
mean (is.na(nactivity))
```

Now we have a first activity which shows our first data base, and the activity database with the na 's filled in.

Here is our new histogram

```{r,}
library(graphics)
hist(tapply(nsteps, ndate ,sum), xlab = "Total daily steps", breaks = 20, main = "Total of steps taken per day")
sum_n <-tapply(nsteps, ndate, sum)
head(sum_n,na.rm = TRUE)
```

If we fill the missing value the data we see will give as less imformation. In this case, there were filled plot squares as na where marked as 0. The 0 value increased a lot making our data noisier.

Now lets subset the data in week days and weekends

```{r,warning=FALSE,R.options=FALSE,echo=TRUE,prompt=FALSE,render=FALSE}
date<- as.Date(date, format = "%Y-%m-%d")
```

```{r,warning=FALSE}
nactivity$day.type <- c("weekend", "weekday", "weekday", "weekday", "weekday", "weekday", "weekend")[as.POSIXlt(nactivity$date)$wday + 1]
nactivity$day.type <- as.factor(nactivity$day.type)
weekday <- nactivity[nactivity$day.type == "weekday", ]
weekend <- nactivity[nactivity$day.type =="weekend", ]
weekday.means <- as.numeric(tapply(weekday$steps, weekday$interval, mean))
weekend.means <- as.numeric(tapply(weekend$steps, weekend$interval, mean))
```

 Here is the plot 
 
```{r}
plot( weekday.means, type = "l", 
     col = "red", ylab = "Average steps", xlab = "Time of day", main = "Comparison between weekday and weekend", 
     xaxt = "n")
lines(weekend.means, type = "l",  col = "blue")
legend(1500, 230, c("Weekend", "Weekday "), lty = c(1, 1), lwd = c(1, 1), col = c("blue", "red"))
```

***Conclusions***

Those who do exercise do it more often during the week, and the sedentary do it in the weekend.


